How did local health infrastructure and socio-political factors within different states and counties in the United States affect the disparities in COVID-19 outcomes, and what lessons can be learned for more targeted public health preparedness and response strategies in future pandemics?

### 2024-02-13 outputs/paper/paper.qmd
data measurements and others

>>>>

The transformation of real-world phenomena into quantifiable data points in our dataset is influenced by various factors, leading to inherent limitations in our  measurements for life expectancy rates, and political affiliation.

Our measurement for COVID death rates is drawn from a accumulated dataset last updated until March 2023. This temporal constraint means we would not be aware of the fluctuations of death rates over time, particularly during the peak of the pandemic in 2020. The data's primarily reliance on the reported statistics from hospitals, clinics, testing centres, and laboratories may introduces bias as excludes uncounted deaths from COVID. Especially during the critical period of health facilities capacity constraints,many individuals died outside of medical facilities without being documented. Therefore, our measurements for COVID death rates are only accounted for the documented death from COVID.

In addition,it's important to note that the death rate presented here reflects the entire population of a state. However, COVID-19 disproportionately affects older individuals compared to younger adults in the prime of their lives. As a result, this data does not capture the effect of COVID across different age groups. Instead, it only offers a general overview the state's population. In summary, we are unable to provide detailed insights into age-specific or temporal variations based on this dataset.

Similarly, our measurement of life expectancy suffer from the demographic limitations. Even though we categorize life expectancy by race, the absence of an age breakdown restricts us discern impacts across different age groups. Moreover, delays in reporting deaths undermine the timeliness and accuracy of our data, potentially skewing our understanding of population health trends.

Our measurement for each state's political preference is calculated from the 2020 election votes. There may be many scenarios that prevented eligible voters from voting. However, we are only able to assess the the state's political preference based on the people who voted, thereby potentially introducing selection bias. Additionally, issues such as ballot loss or voter disenfranchisement further compound the reliability of our measurements.
